---
title: "Evaluation of Omaha, NE Bridge Integrity"
author: "Jack Macfadyen"
format: revealjs
editor: visual
---

## Premise

-   The mayor of Omaha, NE has concerns about degrading transportation infrastructure
-   What is the risk of a bridge collapse event over the next 5 years?
-   What are the risks involved in a bridge collapse event?
-   How can this risk be mitigated?

------------------------------------------------------------------------

## Outline

-   Data and methods used
-   Risk assessment
-   What can be done

------------------------------------------------------------------------

## Data and Methods

**Data Used:** National Bridge Inventory from the U.S. Federal Highway Administration

**Key Variables:**

-   Condition Ratings (0 to 9)
    -   Rating of 0 = failed condition
    -   Rating ≤ 4 = poor condition
    -   Rating \> 4 = fair to good condition
    -   Rating of 9 = excellent condition
-   Year Built
-   Average Daily Traffic

------------------------------------------------------------------------

## Data & Results

**Figure 1: Distribution of bridge condition ratings in Douglas County**

```{r, echo = FALSE}
library(tidyverse)
library(ggplot2)
nbi_data <- read.csv("NE24.txt", header = TRUE)
data <- nbi_data %>%
  filter(COUNTY_CODE_003 == "55")
data %>%
  count(LOWEST_RATING) %>%
  ggplot(aes(x = LOWEST_RATING, y = n)) +
  geom_col(fill = "grey", color = "black") +
  geom_text(aes(label = n, vjust = -0.2)) +
  scale_x_continuous(breaks = unique(sort(data$LOWEST_RATING))) +
  labs(
    x = "Lowest Condition Rating",
    y = "Count"
  ) +
  theme_minimal()
```

------------------------------------------------------------------------

## Data & Results

**Figure 2: Construction year distribution by lowest bridge condition rating**

```{r, echo = FALSE}
ggplot(data, aes(x = YEAR_BUILT_027, y = factor(LOWEST_RATING))) +
  geom_boxplot(fill = "grey", color = "black") +
  labs(
    x = "Year Built",
    y = "Lowest Condition Rating"
  )
```

------------------------------------------------------------------------

## Data & Results

**Figure 3: Average daily traffic by bridge condition rating**

```{r, echo = FALSE}
a <- data %>% 
  filter(LOWEST_RATING == 4) %>% 
  summarise(Avg_Year_Poor = round(mean(YEAR_BUILT_027, na.rm = TRUE), 0)) %>%
  pull(Avg_Year_Poor)

b <- data %>% 
  filter(LOWEST_RATING == 5) %>% 
  summarise(Avg_Year_Not_Poor = round(mean(YEAR_BUILT_027, na.rm = TRUE), 0)) %>% 
  pull(Avg_Year_Not_Poor)

data %>%
  group_by(LOWEST_RATING) %>%
  summarise(mean_ADT = sum(ADT_029, na.rm = TRUE)) %>%
  ggplot(aes(x = LOWEST_RATING, y = mean_ADT)) +
  geom_col(fill = "grey", color = "black") +
  geom_text(aes(label = round(mean_ADT)), vjust = -0.2) +
  scale_x_continuous(breaks = scales::pretty_breaks()) +
  labs(
    x = "Lowest Condition Rating",
    y = "Average Daily Traffic"
  ) +
  theme_minimal()
```

------------------------------------------------------------------------

## Data & Results

**Figure 4: Distribution of average daily traffic by bridge condition rating**

```{r, echo = FALSE}
ggplot(data, aes(x = ADT_029, y = factor(LOWEST_RATING))) +
  geom_boxplot(fill = "grey", color = "black") +
  labs(
    x = "Average Daily Traffic",
    y = "Lowest Condition Rating"
  )
```

------------------------------------------------------------------------

## What can be done

**Priority Repairs:** Target 15 poor-condition bridges

**Traffic Management:** Traffic Management: Monitor high-ADT bridges near I-80

**Funding Strategy:** Allocate budget to replace poor-condition bridges

------------------------------------------------------------------------

## Conclusion

-   **Low collapse risk** but aging infrastructure demands action
-   Current bridge maintenance is successful
-   Sustained funding for long-term safety is vital
